---
title: "Bayesian Analysis of U.S. Support for Military Aid to Ukraine"
author: "Franco Lorenzo"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load all model outputs in order
source("scripts/01_data_setup.R")
source("scripts/02_separate_model.R")
source("scripts/03_pooled_model.R")
source("scripts/04_hierarchical_model.R")
source("scripts/05_predictive_analysis.R")
source("scripts/06_plots_and_summaries.R")
```

# Introduction

This report analyzes data from a February 2025 Economist/YouGov survey of 1,603 U.S. adults regarding their opinions on military aid to Ukraine. It uses Bayesian methods, specifically Gibbs sampling, to fit three models:

- Separate Gaussian model (one mean per group)
- Pooled Gaussian model (single global mean)
- Hierarchical Gaussian model (partial pooling across groups)

Additionally, it predicts the support level for a hypothetical new political group.

The goal is to apply Bayesian inference carefully and discuss the strengths and limitations of these methods in a real-world setting.

# Data Description

The survey asked participants whether the U.S. should:

- Increase Aid
- Maintain Aid
- Decrease Aid
- Are Not Sure

The support percentages by political affiliation are summarized as follows:

| Group        | Increase Aid | Maintain Aid | Not Sure | Decrease Aid |
|--------------|--------------|--------------|----------|--------------|
| Democrat     | 35%           | 39%           | 16%       | 10%           |
| Independent  | 19%           | 23%           | 26%       | 33%           |
| Republican   | 10%           | 24%           | 21%       | 45%           |

Since individual-level responses were not available, I modeled these percentages as observations with Gaussian noise, acknowledging that real-world variance might be underestimated.

# Modeling Approach

The analysis applied Gibbs Sampling across three different model structures:

## Separate Model

Each political affiliation has an independent mean support percentage, sharing a common variance across groups.

```{r separate-model, echo=FALSE, fig.width=12, fig.height=4}
par(mfrow = c(1,3))
for (g in 1:ncol(posterior_mu)) {
  hist(posterior_mu[,g],
       breaks = 30,
       main = paste("Posterior Mean:", colnames(posterior_mu)[g]),
       xlab = "Mean Support (%)",
       col = "lightblue",
       border = "white")
}
```

---

## Pooled Model

All responses are assumed to arise from a single global mean and variance.

```{r pooled-model, echo=FALSE}
hist(posterior_mu_pooled,
     breaks = 30,
     main = "Posterior Distribution (Pooled Model)",
     xlab = "Global Mean Support (%)",
     col = "lightcoral",
     border = "white")
```

---

## Hierarchical Model

The group means are modeled as draws from a common distribution, allowing partial pooling.

```{r hierarchical-model, echo=FALSE, fig.width=12, fig.height=4}
par(mfrow = c(1,3))
for (g in 1:ncol(posterior_mu_g)) {
  hist(posterior_mu_g[,g],
       breaks = 30,
       main = paste("Posterior Mean:", colnames(posterior_mu_g)[g]),
       xlab = "Group Mean (%)",
       col = "lightgreen",
       border = "white")
}
```

---

# Results

## Posterior Summaries

The table below summarizes the posterior mean estimates and 95% credible intervals for each model:

```{r results-plot, echo=FALSE, fig.height=5}
library(ggplot2)

# Create a plot-friendly version of summary_table
plot_data <- summary_table
plot_data$Model <- factor(plot_data$Model, levels = rev(summary_table$Model))  # Reverse for nicer top-down order

# Plot
ggplot(plot_data, aes(x = Mean, y = Model)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbarh(aes(xmin = CI_Lower, xmax = CI_Upper), height = 0.2, color = "gray40") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Posterior Means with 95% Credible Intervals",
    x = "Mean Support (%)",
    y = "Model"
  )
```

---

## Visualizations

Separate Model Posteriors:

```{r separate-histograms, echo=FALSE, fig.width=12, fig.height=4}
par(mfrow = c(1, 3))  # 3 plots in a row
for (g in 1:ncol(posterior_mu)) {
  hist(posterior_mu[, g],
       breaks = 30,
       main = paste("Separate:", colnames(posterior_mu)[g]),
       xlab = "Mean Support (%)",
       col = "lightblue",
       border = "white")
}
```

---

Pooled Model Posterior:

```{r pooled-histogram, echo=FALSE}
hist(posterior_mu_pooled,
     breaks = 30,
     main = "Pooled Model Posterior",
     xlab = "Global Mean Support (%)",
     col = "lightcoral",
     border = "white")
```

---

Hierarchical Model Posteriors:

```{r hierarchical-histograms, echo=FALSE, fig.width=12, fig.height=4}
par(mfrow = c(1, 3))  # layout: 3 plots in one row

for (g in 1:ncol(posterior_mu_g)) {
  hist(posterior_mu_g[, g],
       breaks = 30,
       main = paste("Hierarchical:", colnames(posterior_mu_g)[g]),
       xlab = "Group Mean (%)",
       col = "lightgreen",
       border = "white")
}
```

---

Predictive Distribution (New Group):

```{r predictive-histogram, echo=FALSE}
hist(predictive_mu_new,
     breaks = 30,
     main = "Predictive Distribution: New Group",
     xlab = "Predicted Mean Support (%)",
     col = "lightyellow",
     border = "gray40")
```

---

# Predictive Analysis

From the hierarchical model, the report derives a predictive distribution for a hypothetical new political affiliation's mean support level.

The posterior mean and 95% credible interval for the new group's support level are:

```{r predictive-summary, echo=FALSE, fig.height=4, fig.width=7}
# Compute mean and interval
pred_mean <- mean(predictive_mu_new)
pred_ci <- quantile(predictive_mu_new, probs = c(0.025, 0.975))

# Plot
hist(predictive_mu_new,
     breaks = 30,
     col = "lightyellow",
     border = "gray40",
     main = "Predictive Mean Support for New Group",
     xlab = "Mean Support (%)")

# Add lines for mean and 95% CI
abline(v = pred_mean, col = "blue", lwd = 2)
abline(v = pred_ci, col = "red", lty = 2, lwd = 2)
legend("topright",
       legend = c("Posterior Mean", "95% Credible Interval"),
       col = c("blue", "red"),
       lty = c(1, 2), lwd = 2, bty = "n")
```

---

# Discussion

## Conclusions from the Analysis

- Democrats show the highest mean support for maintaining or increasing aid to Ukraine.  
- Independents are more divided, with a noticeable share unsure or favoring decrease.  
- Republicans show stronger support for decreasing aid compared to other affiliations.  

The hierarchical model offers a **balanced** view by borrowing strength across groups, improving estimates especially where sample sizes are small.

---

## Suitability of the Methods

Bayesian hierarchical modeling is **highly suitable** for this situation:

- It allows partial pooling.  
- It quantifies uncertainty.  
- It can predict for unobserved groups.  

However, care must be taken because:

- The small survey size and absence of raw responses reduce robustness.  
- Real-world data might have non-Gaussian features (e.g., skewness).  

---

## Trust and Limitations

- The model assumes normality and constant variance, which may not perfectly match real opinion distributions.  
- Lack of individual-level data introduces unaccounted variability.  
- Small group sample sizes make inference riskier.  

Still, Bayesian hierarchical models offer the **best available structured estimation** under these limitations.
